\documentclass{article}
\usepackage{amsmath, amssymb, amsthm, graphicx, hyperref, xcolor, longtable}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{natbib}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

\title{\textbf{Cayley--Dickson Structure Census: \\
A Computational Verification Report}\\[0.5em]
\large From Sedenion Box-Kites to Higher-Dimensional Scaling Laws}
\author{open\_gororoba Project}
\date{February 2026}

\begin{document}

\maketitle

\begin{abstract}
We report computational results from the \texttt{open\_gororoba} research
workbench concerning the algebraic structure of Cayley--Dickson algebras at
successive doublings.  All results are organized according to a verification
ladder separating \textbf{verified algebra} (deterministic tests against known
results), \textbf{statistical findings} (permutation-tested hypotheses), and
\textbf{speculative models} (clearly marked).  Each claim references a specific
entry in the project's claims evidence matrix (C-nnn) and the corresponding
Rust test function.
\end{abstract}

\tableofcontents
\newpage

%% ===================================================================
\section{Cayley--Dickson Construction}
\label{sec:cd}

The Cayley--Dickson doubling process \cite{baez2002} constructs a tower of
$2^n$-dimensional real algebras:
\begin{equation}
  A_{n+1} = A_n \oplus A_n, \quad
  (a,b)(c,d) = (ac - \bar{d}b,\; da + b\bar{c})
\end{equation}
yielding the reals ($n=0$), complex numbers ($n=1$), quaternions ($n=2$),
octonions ($n=3$), sedenions ($n=4$), pathions ($n=5$), and beyond.

At each doubling, algebraic properties degrade:
\begin{itemize}
  \item $n=1$: ordering lost
  \item $n=2$: commutativity lost
  \item $n=3$: associativity lost
  \item $n=4$: alternative identity and norm-composition lost; zero divisors appear
  \item $n=5$: power-associativity degrades
\end{itemize}

Our implementation uses integer-exact arithmetic for basis-element products,
avoiding floating-point error entirely (claims C-001 through C-006, verified
by \texttt{algebra\_core::cayley\_dickson} tests).

%% ===================================================================
\section{Zero-Divisor Census and Box-Kite Structure}
\label{sec:zd}

At $\dim \geq 16$ (sedenions), zero divisors organize into combinatorial
structures first described by de~Marrais \cite{demarrais2000, demarrais2004boxkites}.

\subsection{Sedenion Box-Kites}

The 42 primitive cross-assessors of the sedenion algebra partition into 7
box-kites, each containing 6 assessors arranged as 3 ``strut'' pairs.
De~Marrais's published strut table matches our computational census exactly
(claim C-454, test \texttt{test\_strut\_table\_matches\_de\_marrais\_flying\_higher}).

Each box-kite's assessor interaction graph is the complete tripartite graph
$K_{2,2,2}$ (octahedron).  This structure is verified in
\texttt{algebra\_core::boxkites} (claims C-050 through C-060).

\subsection{Cross-Assessor Pairs and the XOR Condition}

The zero-divisor product condition for 2-blade cross-assessor pairs admits a
necessary (but not sufficient) XOR filter: for a product $\alpha\beta = 0$
to hold, the basis indices must satisfy specific GF(2)-linear conditions.
The XOR filter passes 315/168 pairs (ratio 0.533) with zero false negatives
(claim C-445, insight I-012).

%% ===================================================================
\section{Motif Scaling Laws}
\label{sec:motifs}

Exact enumeration of the zero-product graph's connected components across
5 Cayley--Dickson doublings ($\dim = 16, 32, 64, 128, 256$) reveals strict
scaling laws (insight I-006, claims C-126 through C-130):

\begin{center}
\begin{tabular}{lcc}
\hline
\textbf{Invariant} & \textbf{Formula} & \textbf{Verified dims} \\
\hline
Components & $\dim/2 - 1$ & 16--256 \\
Nodes per component & $\dim/2 - 2$ & 16--256 \\
Motif classes & $\dim/16$ & 16--256 \\
$K_2$ components & $3 + \log_2(\dim)$ & 16--256 \\
$K_2$ part count & $\dim/4 - 1$ & 16--256 \\
\hline
\end{tabular}
\end{center}

No octahedra ($K_{2,2,2}$) appear beyond $\dim = 16$.  At $\dim = 32$, the
structure completely reorganizes into 8 heptacross ($K_{2,2,2,2,2,2,2}$) and
7 mixed-degree components (insight I-006).

All motif census computations are exact (not sampled) and complete in under
2 seconds for $\dim = 256$ in release mode.

%% ===================================================================
\section{Lattice Codebook Filtration}
\label{sec:lattice}

Cayley--Dickson basis elements at $\dim = 256, 512, 1024, 2048$ embed into an
8-dimensional integer lattice with coordinates in $\{-1, 0, 1\}$
(claims C-452, C-453, insight I-014).  This embedding is dimension-independent:
all four algebras use the same 8D target, suggesting the octonion sub-algebra
provides the fundamental lattice structure.

Eight monograph theses (A--H) verified \cite{reggiani2024} (insight I-015):

\begin{description}
  \item[Thesis A] Codebook parity: coordinates sum to even, nonzero count is even,
    first coordinate $\neq +1$ (C-458).
  \item[Thesis B] Filtration nesting: $\Lambda_{256} \subset \Lambda_{512}
    \subset \Lambda_{1024} \subset \Lambda_{2048}$ (C-459).
  \item[Thesis C] Prefix-cut transitions: each child is a lexicographic prefix
    of its parent (C-460).
  \item[Thesis D] Scalar shadow invariant (C-461).
  \item[Thesis E] XOR partner law: $\text{partner}(i) = i \oplus (N/16)$ for
    $N \geq 64$ (C-462).
  \item[Thesis F] Parity-clique decomposition: $K_4 \cup K_4$ at $\dim = 16$,
    $K_8 \cup K_8$ at $\dim = 32$ (C-463).
  \item[Thesis G] Spectral fingerprints: $S_{\text{base}} = 2187 = 3^7$ (C-464).
  \item[Thesis H] Null-model rejection (C-465).
\end{description}

The E8 root-system connection was tested and \textbf{refuted}: zero out of 336
pairwise lattice differences have norm-squared 2 (claim C-455, insight I-014).

%% ===================================================================
\section{Emanation Architecture}
\label{sec:emanation}

The de~Marrais emanation construction \cite{demarrais2006} builds 18 layers
(L1--L18) from the sedenion box-kite structure.  Our implementation verifies
(insight I-016, claims C-468 through C-475):

\begin{itemize}
  \item DMZ = sign-concordance (12 edges per box-kite)
  \item Sail-loop = Fano incidence pattern
  \item Oriented Trip Sync is universal across all 7 box-kites
  \item 113 dedicated tests, 4400 lines of implementation
\end{itemize}

%% ===================================================================
\section{Ultrametric Hierarchy Fingerprint}
\label{sec:ultrametric}

Statistical analysis across 9 astrophysical catalogs using GPU-accelerated
ultrametric fraction testing (10 million triples, 1000 permutations, RTX~4070~Ti)
reveals hierarchical structure in specific physical regimes
(insight I-011, I-013; claims C-436 through C-444):

\begin{center}
\begin{tabular}{lccc}
\hline
\textbf{Catalog} & \textbf{Significant tests} & \textbf{p range} & \textbf{Verdict} \\
\hline
Hipparcos & 48/114 & $< 0.05$ & galactic hierarchy \\
CHIME/FRB & 8/8 & $< 0.01$ & ISM-mediated DM \\
ATNF Pulsars & 6/8 & $< 0.01$ & ISM-mediated DM \\
GWOSC GW & 4/66 & $< 0.01$ & mass-ratio clustering \\
SDSS Quasars & 2/52 & marginal & weak \\
Pantheon+ SNe & 2/22 & marginal & light-curve shape \\
McGill Magnetars & 0/66 & -- & $N$ too small \\
Fermi GBM GRBs & 0/22 & -- & isotropic \\
\hline
\end{tabular}
\end{center}

The test discriminates physical hierarchy from noise: catalogs with known
hierarchical organization (Galactic proper motions, ISM electron-density
structure) show strong signal; isotropic catalogs show none (insight I-013).

%% ===================================================================
\section{Physical Models (Speculative Tier)}
\label{sec:speculative}

\textbf{Status: speculative.}  The following models are exploratory and have
not been validated against observational data.  They are tracked in the claims
evidence matrix with explicit falsification criteria.

\begin{description}
  \item[Gravastar TOV] Three-layer gravastar parameter sweep across polytropic
    indices (claims C-400 through C-410).  Numerical solutions exist but physical
    relevance is untested.
  \item[Bounce cosmology] Joint Pantheon+/DESI fit yields $\Delta\text{BIC} = +7.37$
    favoring $\Lambda$CDM over bounce model (claims C-200 through C-210,
    insight I-004).  The quantum correction parameter converges to zero.
  \item[Negative-dimension eigenvalues] Fractional Laplacian eigenvalue
    convergence under $\epsilon \to 0$ (claims C-420 through C-425).
\end{description}

%% ===================================================================
\appendix

\input{claims_appendix}
\input{insights_appendix}

%% ===================================================================
\bibliographystyle{plainnat}
\bibliography{cayley_dickson}

\end{document}
