\documentclass{article}
\usepackage{amsmath, amssymb, amsthm, graphicx, hyperref}
\usepackage[a4paper, margin=1in]{geometry}

\title{Mathematical Formalism of Cayley--Dickson Zero-Divisor Structure}
\author{open\_gororoba Project}
\date{February 2026}

\begin{document}

\maketitle

\section{Scope and Validation Status}

This document describes the mathematical formalism underlying the
\texttt{open\_gororoba} codebase.  It separates:
\begin{itemize}
  \item standard definitions (Cayley--Dickson construction, associator),
  \item computationally validated facts (verified by Rust unit tests in
    \texttt{crates/algebra\_core/}), and
  \item speculative proposals that are clearly marked as such.
\end{itemize}

Treat any statement as non-authoritative unless backed by a first-party
citation and a reproducible test under \texttt{crates/*/src/} or the claims
evidence matrix (\texttt{registry/claims.toml}).

\section{Cayley--Dickson Construction (Standard)}

\subsection{Recursive Definition}
Let $A_n$ be the algebra obtained after $n$ iterations of the Cayley--Dickson
process, with $\dim(A_n) = 2^n$.  The standard doubling is:
\begin{equation}
  A_{n+1} = A_n \oplus A_n, \quad
  (a,b)(c,d) = (ac - \bar{d}b,\; da + b\bar{c})
\end{equation}

\section{Trilinear Order-Sensitivity Operators}

\subsection{Associator (Standard)}
For octonions and beyond, associativity fails.  The standard associator is:
\begin{equation}
  [x,y,z] = (xy)z - x(yz)
\end{equation}

\subsection{The \texorpdfstring{$m_3$}{m3} Operator (Validated Properties)}

The repo computes a specific trilinear operator $m_3$ (defined in
\texttt{crates/algebra\_core/src/m3\_transfer.rs}).  On distinct octonion
basis triples $(e_i, e_j, e_k)$ with $i,j,k \in \{1,\dots,7\}$ all distinct,
the following properties are validated:
\begin{itemize}
  \item exactly 42 scalar outputs and 168 pure-imaginary outputs, and
  \item the scalar cases correspond to the Fano-plane lines (with parity
    sign flips).
\end{itemize}
These are validated by \texttt{crates/algebra\_core/src/m3\_transfer.rs}
test module.  Establishing an $A_\infty$ structure would require additional,
separately scoped Stasheff-identity checks.

\section{Sedenion Zero Divisors}

\subsection{The 42 Assessors}
In the de~Marrais construction, there is a set of 42 primitive cross-assessors
that generate a diagonal-form family of 2-blades.  This repo replicates:
\begin{itemize}
  \item 42 primitive assessors, and
  \item a partition into 7 box-kites (6 assessors each)
\end{itemize}
under the repo's fixed 16D Cayley--Dickson convention, validated by
\texttt{crates/algebra\_core/src/boxkites.rs} test module.

\subsection{Box-Kite Structure}
\emph{Box-kite} is a specific combinatorial object: the complete tripartite
graph $K_{2,2,2}$.  At $\dim = 16$, all 7 connected components of the
cross-assessor zero-product graph are isomorphic to $K_{2,2,2}$.

\section{Results Established Since January 2026}

The following items were listed as ``speculative directions'' in the original
January 2026 draft.  Their status has been resolved:

\begin{description}
  \item[Motif census (dim=32/64/128/256)] \textbf{COMPLETED} (Sprint 6).
    Exact scaling laws verified across 5 doublings.  No octahedra beyond
    $\dim = 16$; structure completely reorganizes at each doubling.
    See \texttt{crates/algebra\_core/src/boxkites.rs}, claims C-126 through C-130.

  \item[Reggiani geometric identifications] \textbf{PARTIALLY COMPLETED}
    (Sprint 9).  8 monograph theses (A--H) verified, including lattice
    codebook filtration and parity constraints.  Full geometric mapping
    (PSL(2,7) action) remains open.

  \item[E8 root-system mapping] \textbf{REFUTED} (Sprint 8).  Zero out of 336
    pairwise lattice-point differences have norm-squared 2.  The minimum
    lattice separation for ZD-adjacent pairs is $|d|^2 = 4$.  See claim C-455.
\end{description}

\section{Open Questions}

\begin{itemize}
  \item \textbf{Twist-delta pair correspondence} (C-466): relationship between
    de~Marrais twist pairs and lattice codebook structure.
  \item \textbf{PSL(2,7) geometric mapping}: explicit construction of the
    projective special linear group action on box-kite coordinates.
  \item \textbf{Higher-dimension emanation}: extending L1--L18 emanation
    layers beyond $\dim = 16$ to $\dim = 32$ and beyond.
\end{itemize}

\end{document}
